---
title: "Untitled"
author: "Maria Olsen"
date: "2022-10-05"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
pacman::p_load(tidyverse,brms)
```

#The workflow (from Riccardo)

- define the formula
- define the prior
- prior predictive checks
- fit the model
- model quality checks: traceplots, divergences, rhat, effective samples #presition analysis
- model quality checks: posterior predictive checks, prior-posterior update checks
- model comparison #cross validation

#Simulating data
```{r}
set.seed(1912)

#Defining parameters 
n <- 30 #set to 50?
mu_asd <- log(1.5)
sigma_asd <- log(1.5)-log(1.5-0.5)
mu_td <- log(1.5)
sigma_td <- log(1.5)-log(1.5-0.3)

mu_visit_asd <- 0.15  
sigma_visit_asd <-  0.1 

mu_visit_td <-  0.20 
sigma_visit_td <- 0.08  

visit <- 6
error <- 0.2

#Making a function for simulating data
s_d <- function(n, visit, mu_asd, mu_td, sigma_asd, sigma_td, error){
  s_df <- tibble(expand.grid(ID=seq(n),
                             Diag= c("ASD", "TD"),
                             Visit = seq(visit))) %>%  
    mutate(ID = ifelse(Diag == "TD", ID + (n*2), ID), 
           IndividualIntercept = NA, 
           IndividualSlope = NA, 
           MLU = NA)
  
  for (i in seq(s_df$ID)) {
    #Assigning individual intercept
    s_df$IndividualIntercept[s_df$ID == i & s_df$Diag == "ASD"] <- rnorm(1, mu_asd, sigma_asd)
    s_df$IndividualIntercept[s_df$ID == i & s_df$Diag == "TD"] <- rnorm(1, mu_td, sigma_td)
    
    #Assigning individual slope
    s_df$IndividualSlope[s_df$ID == i & s_df$Diag == "ASD"] <- rnorm(1, mu_visit_asd, sigma_visit_asd)
    s_df$IndividualSlope[s_df$ID == i & s_df$Diag == "TD"] <- rnorm(1, mu_visit_td, sigma_visit_td)
  }
  
  for (i in seq(nrow(s_df))){
  s_df$MLU[i] <- exp(rnorm(1, (s_df$IndividualIntercept[i] + s_df$IndividualSlope[i] * (s_df$Visit[i]-1)), error))
                  }
  
  
  
  return(s_df)
}

d <- s_d(n, visit, mu_asd, mu_td, sigma_asd, sigma_td, error)

#Visualizing data
ggplot(d, aes(Visit, MLU, color = Diag, group = ID)) + 
  theme_bw() + 
  geom_point() + 
  geom_line(alpha = 0.3)

```


#Defining the formula
```{r}
#Intercepts only model
MLU_f0 <- bf(MLU ~ 1)

#Find out what these models are!!
MLU_f1 <- bf(MLU ~ 0 + Diag)

#Interceot + slope
MLU_f2 <- bf(MLU ~ 0 + Diag + Diag:Visit)

#Interept + slope and varrying intercept and slope
MLU_f_3 <- bf(MLU ~ 0 + Diag + Diag:Visit
+ (1 + Visit|ID))
```

#Making priors and prior predictive check
```{r}
#Looking parameters for prior
get_prior(MLU_f1, 
          data = d, 
          family = lognormal)


get_prior(MLU_f2,
          data = d,
          family = lognormal)


get_prior(MLU_f_3,
          data = d,
          family = lognormal)
 


#making priors
MLU_f1_prior <- c(
  prior(normal(0.41, 0.41), class=b, coef= "DiagASD"),
  prior(normal(0.41, 0.22), class=b, coef= "DiagTD"),
  prior(normal(0, 2), class= sigma)
)

MLU_f2_prior <- c(
  prior(normal(0, 0.2), class=b, lb=0), #error and lb=lower boundries
  prior(normal(0.41, 0.41), class=b, coef= "DiagASD"),
  prior(normal(0.41, 0.22), class=b, coef= "DiagTD"),
  prior(normal(0.15, 0.1), class=b, coef= "DiagASD:Visit"),
  prior(normal(0.2, 0.08), class=b, coef= "DiagTD:Visit"),
  prior(normal(0, 0.2), class= sigma)
)

MLU_f3_prior <- c(
  prior(normal(0, 0.2), class=b, lb=0),
  prior(normal(0.41, 0.41), class=b, coef= "DiagASD"),
  prior(normal(0.41, 0.22), class=b, coef= "DiagTD"),
  prior(normal(0.15, 0.1), class=b, coef= "DiagASD:Visit"),
  prior(normal(0.2, 0.08), class=b, coef= "DiagTD:Visit"),
  prior(normal(0, 0.2), class=sd, coef= Intercept, group=ID), #allowing the intercept for each person to varriate with 40% (because of logscale)
  prior(normal(0, 0.1), class=sd, coef= Visit, group=ID), #slope to varriate with 20% for each person
  prior(normal(0, 0.2), class= sigma),
  prior(lkj(1), class= "cor") 
)

MLU_f1_prior_samp <- 
  brm(
    MLU_f1, 
    data = d,
    family = lognormal,
    prior = MLU_f1_prior,  
    sample_prior = "only", 
    iter = 2000,
    warmup = 500,
    backend = "cmdstanr",
    threads = threading(2),
    cores = 2,
    chains = 2,
    file = "MLU_f1_prior_samp",
    control = list(adapt_delta = 0.99, max_treedepth = 20))


MLU_f2_prior_samp <- 
  brm(
    MLU_f2, 
    data = d,
    family = lognormal,
    prior = MLU_f2_prior,  
    sample_prior = "only", 
    iter = 2000,
    warmup = 500,
    backend = "cmdstanr",
    threads = threading(2),
    cores = 2,
    chains = 2,
    file = "MLU_f2_prior_samp",
    control = list(adapt_delta = 0.99, max_treedepth = 20))


MLU_f3_prior_sam <- 
  brm(
    MLU_f_3, 
    data = d,
    family = lognormal,
    prior = MLU_f3_prior,  
    sample_prior = "only", 
    iter = 2000,
    warmup = 500,
    backend = "cmdstanr",
    threads = threading(2),
    cores = 2,
    chains = 2,
    file = "MLU_f3_prior_sam",
    control = list(adapt_delta = 0.99, max_treedepth = 20))


#pp checking the priors 
pp_check(MLU_f1_prior_samp, ndraws = 100)
pp_check(MLU_f2_prior_samp, ndraws = 100)
pp_check(MLU_f3_prior_sam, ndraws = 100)

```


#fitting the models and posterior predictive check
```{r}
#Making posteriors
MLU_f1_prior_posterior <- 
  brm(
    MLU_f1, 
    data = d,
    family = lognormal,
    prior = MLU_f1_prior,  
    sample_prior = T, 
    iter = 5000,
    warmup = 1000,
    backend = "cmdstanr",
    threads = threading(2),
    cores = 2,
    chains = 2,
    file = "MLU_f1_prior_posterior",
    control = list(adapt_delta = 0.99, max_treedepth = 20))


MLU_f2_prior_posterior <- 
  brm(
    MLU_f2, 
    data = d,
    family = lognormal,
    prior = MLU_f2_prior,  
    sample_prior = T, 
    iter = 5000,
    warmup = 1000,
    backend = "cmdstanr",
    threads = threading(2),
    cores = 2,
    chains = 2,
    file = "MLU_f2_prior_posterior",
    control = list(adapt_delta = 0.99, max_treedepth = 20))


MLU_f3_prior_posterior <- 
  brm(
    MLU_f_3, 
    data = d,
    family = lognormal,
    prior = MLU_f3_prior,  
    sample_prior = T, 
    iter = 5000,
    warmup = 1000,
    backend = "cmdstanr",
    threads = threading(2),
    cores = 2,
    chains = 2,
    file = "MLU_f3_prior_posterior",
    control = list(adapt_delta = 0.99, max_treedepth = 20))

#Checking them
p1 <- pp_check(MLU_f1_prior_posterior, ndraws = 100)+ggtitle('Model 1')
p2 <- pp_check(MLU_f2_prior_posterior, ndraws = 100)+ggtitle('Model 2')
p3 <- pp_check(MLU_f3_prior_posterior, ndraws = 100)+ggtitle('Model 3')
gridExtra::grid.arrange(p1,p2,p3)
```


#prior-posterior update checks
```{r}

#getting an overview of the parameters in posterior
variables(MLU_f1_prior_posterior)
variables(MLU_f2_prior_posterior)
variables(MLU_f3_prior_posterior)

#sampling from model and storing it (so we can use it to visualize prior-posterior update check/plot)
MLU_f1_pp_samp <- as_draws_df(MLU_f1_prior_posterior)
MLU_f2_pp_samp <- as_draws_df(MLU_f2_prior_posterior)
MLU_f3_pp_samp <- as_draws_df(MLU_f3_prior_posterior)


#################here my knowledge ends. But example of how to make prior-posterior uddate plot, just need the right varriable names



#Plot the prior-posterior update plot for the intercept:
ggplot(MLU_f1_pp_samp) +
  geom_density(aes(prior_Intercept), fill="steelblue", color="black",alpha=0.6) +
  geom_density(aes(b_Intercept), fill="#FC4E07", color="black",alpha=0.6) + 
  theme_classic()
#try coloring by group to see if it works

#Plot the prior-posterior update plot for the sigma:
ggplot(Posterior_f0) +
  geom_density(aes(prior_sigma), fill="steelblue", color="black",alpha=0.6) +
  geom_density(aes(sigma), fill="#FC4E07", color="black",alpha=0.6) + 
  theme_classic()
```

```{r}
plot(conditional_effects(mlu1p1fit), points = T)
summary(MLU_f1_prior_posterior
        )

MLU_f1_prior_posterior, ndraws = 100)+ggtitle('Model 1')
p2 <- pp_check(MLU_f2_prior_posterior, ndraws = 100)+ggtitle('Model 2')
p3 <- pp_check(MLU_f3_prior_posterior

# model 2
plot(conditional_effects(mlu2p2fit), points = T)
summary(MLU_f2_prior_posterior)

#model 3 Parts of the model have not converged (some Rhats are > 1.05). Be careful when analysing the results! We recommend running more iterations and/or setting stronger priors
plot(conditional_effects(mlu3p3fit), points = T)
summary(MLU_f3_prior_posterior)
```
Family: lognormal 
  Links: mu = identity; sigma = identity 
Formula: MLU ~ 0 + Diag 
   Data: d (Number of observations: 360) 
  Draws: 2 chains, each with iter = 5000; warmup = 1000; thin = 1;
         total post-warmup draws = 8000

Population-Level Effects: 
        Estimate Est.Error l-95% CI u-95% CI Rhat Bulk_ESS Tail_ESS
DiagASD     0.66      0.04     0.58     0.74 1.00     5972     4445
DiagTD      0.92      0.04     0.84     1.00 1.00     6013     5009

Family Specific Parameters: 
      Estimate Est.Error l-95% CI u-95% CI Rhat Bulk_ESS Tail_ESS
sigma     0.56      0.02     0.52     0.60 1.00     6055     5151


 Family: lognormal 
  Links: mu = identity; sigma = identity 
Formula: MLU ~ 0 + Diag + Diag:Visit 
   Data: d (Number of observations: 360) 
  Draws: 2 chains, each with iter = 5000; warmup = 1000; thin = 1;
         total post-warmup draws = 8000

Population-Level Effects: 
              Estimate Est.Error l-95% CI u-95% CI Rhat Bulk_ESS Tail_ESS
DiagASD           0.16      0.07     0.02     0.31 1.00     2630     1917
DiagTD            0.27      0.07     0.13     0.41 1.00     3318     2960
DiagASD:Visit     0.14      0.02     0.10     0.18 1.00     2734     3420
DiagTD:Visit      0.19      0.02     0.15     0.23 1.00     3348     3146

Family Specific Parameters: 
      Estimate Est.Error l-95% CI u-95% CI Rhat Bulk_ESS Tail_ESS
sigma     0.47      0.02     0.43     0.50 1.00     4684     4391

Family: lognormal 
  Links: mu = identity; sigma = identity 
Formula: MLU ~ 0 + Diag + Diag:Visit + (1 + Visit | ID) 
   Data: d (Number of observations: 360) 
  Draws: 2 chains, each with iter = 5000; warmup = 1000; thin = 1;
         total post-warmup draws = 8000

Group-Level Effects: 
~ID (Number of levels: 60) 
                     Estimate Est.Error l-95% CI u-95% CI Rhat Bulk_ESS Tail_ESS
sd(Intercept)            0.30      0.04     0.23     0.38 1.00     4218     5566
sd(Visit)                0.10      0.01     0.08     0.13 1.00     1887     3014
cor(Intercept,Visit)    -0.25      0.15    -0.52     0.07 1.00     1219     2213

Population-Level Effects: 
              Estimate Est.Error l-95% CI u-95% CI Rhat Bulk_ESS Tail_ESS
DiagASD           0.15      0.06     0.03     0.28 1.00     3467     3257
DiagTD            0.27      0.06     0.14     0.39 1.00     3840     3813
DiagASD:Visit     0.15      0.02     0.11     0.18 1.00     2449     3932
DiagTD:Visit      0.19      0.02     0.16     0.23 1.00     2836     3923

Family Specific Parameters: 
      Estimate Est.Error l-95% CI u-95% CI Rhat Bulk_ESS Tail_ESS
sigma     0.20      0.01     0.19     0.22 1.00     5598     6092

Draws were sampled using sample(hmc). For each parameter, Bulk_ESS
and Tail_ESS are effective sample size measures, and Rhat is the potential
scale reduction factor on split chains (at convergence, Rhat = 1).



Draws were sampled using sample(hmc). For each parameter, Bulk_ESS
and Tail_ESS are effective sample size measures, and Rhat is the potential
scale reduction factor on split chains (at convergence, Rhat = 1).

Draws were sampled using sample(hmc). For each parameter, Bulk_ESS
and Tail_ESS are effective sample size measures, and Rhat is the potential
scale reduction factor on split chains (at convergence, Rhat = 1).

#Notes for how I understand the process so far
We are simulating data that we think/are trying to recreate our real data so we can have an idea of how the real data would work and look like when we start analyzing it (so analysing the fake data and hope that it gives an estimate of how the real data would work).

We start by making the models we want to use (and want to find the best one of these models, so we can just use that one model when we get to analyzing the real data). 

Then we make priors based on the values from our simulation and check them - they look pretty wierd, but that's ok I think.

Then we make posteriors based on our priors and simulated data, and check them - they should look much better.

Then we sample from the posterior so we can visualize if our simulated data has learned from our priors, and how well it has learned. 
I think making a summery would be a good idea, but "testing our hypothesis" by using the hypothesis(), so we can look at the Post.Prob value, that should be as close to 1 as posible. The model with the best Post.Prob value should be the one we use when we analyze the real data, as well as the priors used with that model.




#Hilsen fra Maria
H책ber at det giver ok mening, det er i hvert fald alt hvad jeg selv forst책r indtil videre - h책ber at du f책r R til at virke! 



